---
title: "GallupDataPreparation"
author: "Kenith"
date: '2023-08-28'
---

## 0. Subset the Gallup raw data

The version of the Gallup raw data used in this study is "The_Gallup_021723.dat". This data was released to the [UofT Map and Data Library](https://mdl.library.utoronto.ca/collections/numeric-data/gallup) on February 14, 2023.

Since the raw data file is relatively large (14Gb), we used a separate environment to run the following code, which subsets the variables that we needed.

```{r eval=FALSE}
# This code will not be run
# Import the data file "The_Gallup_021723.dat"
gallupRaw <- rio::import("The_Gallup_021723.dat")

# Create a character vector of the names of the variables we're interested in
varRequired <- c(
    "WPID", "WGT", "COUNTRYNEW", "COUNTRY_ISO3", "YEAR_CALENDAR", "WP4", "YEAR_WAVE", 
    "WP16"
)

# Subset gallupRaw to include only the variables specified in varRequired
gallupSubset <- gallupRaw %>% select(all_of(varRequired))

# Save the gallupSubset data frame to a .rds file for later use
saveRDS(gallupSubset, "GWP_subset_ethics.rds")
```

The environment used had the following specifications:\
R version 4.0.3 (2020-10-10)

Platform: x86_64-pc-linux-gnu (64-bit)

Running under: CentOS Linux 7 (Core)

-   **R version**: 4.0.3 (2020-10-10)

-   **Platform**: x86_64-pc-linux-gnu (64-bit)

-   **Running under**: CentOS Linux 7 (Core)

-   **Matrix products**: default

    -   **BLAS**: /local/software/R/v4.0.3/lib64/R/lib/libRblas.so
    -   **LAPACK**: /local/software/R/v4.0.3/lib64/R/lib/libRlapack.so

-   **Attached base packages**: stats, graphics, grDevices, utils, datasets, methods, base

-   **Other attached packages**: dplyr_1.0.6

-   **Loaded via a namespace (and not attached)**: base64enc_0.1-3, bit_4.0.4, bit64_4.0.5, cellranger_1.1.0, compiler_4.0.3, crayon_1.4.1, curl_4.3.1, data.table_1.14.0, digest_0.6.27, ellipsis_0.3.2, evaluate_0.14, fansi_0.5.0, forcats_0.5.1, foreign_0.8-80, generics_0.1.0, glue_1.4.2, haven_2.4.1, hms_1.1.0, htmltools_0.5.1.1, IRdisplay_1.0, IRkernel_1.2, jsonlite_1.7.2, lifecycle_1.0.0, magrittr_2.0.1, openxlsx_4.2.3, pbdZMQ_0.3-5, pillar_1.6.1, pkgconfig_2.0.3, purrr_0.3.4, R6_2.5.0, Rcpp_1.0.6, readxl_1.3.1, repr_1.1.3, rio_0.5.26, rlang_0.4.11, stringi_1.6.2, tibble_3.1.2, tidyselect_1.1.1, tools_4.0.3, utf8_1.2.1, uuid_0.1-4, vctrs_0.3.8, zip_2.2.0

## 1. Clean the Subset Gallup Data

Once we have the subset of the Gallup data, we import the data and load the necessary packages:

```{r}
# Load "dplyr"
library(dplyr)

# Read the RDS file
gallupSubset <- readRDS("GWP_subset_ethics.rds")
```

Next, we create a new variable, `YEAR_INTERVIEW`, which is the year of the interview. This is either: - Extracted from `WP4` (Interview Date), or - The same as `YEAR_CALENDAR` (The calendar year during which the data were collected, based on the last day of interviewing) if `WP4` is missing.

```{r}
gallup <- gallupSubset %>% 
  # Convert WP4 to a date
  mutate(WP4 = as.Date(WP4, "%m/%d/%Y")) %>%
  # Create YEAR_INTERVIEW variable
  mutate(YEAR_INTERVIEW = ifelse(is.na(WP4), YEAR_CALENDAR, format(WP4, format="%Y")))
```

Gallup has coded some responses as 'Don't Know' or 'Refused'. These responses are not meaningful for our analysis, so we'll convert them to `NA`. The corresponding values can be found at the [WORLD POLL REFERENCE TOOL](https://wpr.gallup.com/).

```{r}
# Convert certain values to NA
gallup <- gallup %>% mutate(WP16 = ifelse(WP16 %in% c(98, 99), NA, WP16))
```

The object `gallup` is now a cleaned subset of the Gallup data.
